---
title: "Ozias sensitivity"
author: "JP"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data and packages

```{r}
# packages
packages <- c("here", "tidyverse", "readxl", "RColorBrewer", "scico")

# install if not installed
not_installed <- packages[!(packages %in% installed.packages()[,"Package"])]

if (length(not_installed) > 0) {
  install.packages(not_installed)
}

# Load packages
lapply(packages, library, character.only = TRUE)


# After once installed, this is faster:
library(here)
library(tidyverse)
library(readxl)
library(RColorBrewer)
library(scico)  

# Exact colours used:
scico(4, palette = "managua")
exact_colours <- c("#FFCE66", "#92463A", "#4D5492", "#80E6FF")


# Convert ES to perc
perc<- function(data){
  data_conv=100*(exp(data) - 1)
  return(data_conv)
}

# Read the data
df_raw <- read_excel("dataset_Ozias.xlsx")



# Explore data: number of studies and crops
length(unique(df_raw$key)) # 4 studies?
length(unique(df_raw$Crop_Group)) # 8 crop

names(df_raw)

```


```{r}
# Group data by studies and count the number of cases
df_raw<-df_raw %>% 
  mutate(ES=perc(effectSize)) %>%
  filter(!ES > 100) %>%
  dplyr::select(!effectSize)  %>%
  dplyr::rename(effectSize=ES) 

  
 df_raw %>% group_by(key) %>% 
  summarise(Number_of_cases = n()) %>% 
  arrange(Number_of_cases) # OF 587, AF 853, CC 1029, NT 7763
```

```{r}
tail(df_raw)
```




```{r}
# Calculate quantiles outside ggplot for cleaner code
lower_quantile <- df_raw |> pull(effectSize) |> quantile(0.025)
upper_quantile <- df_raw |> pull(effectSize) |> quantile(0.975)
```





```{r}

```

# Plot distributions for each key - density

EDIT September 2024:
No need to separate crops: crops within the same key can be combined
Then, for each key calculate how much the mean value (within a key) changes if we drop one observation at a time

```{r}
df_raw %>%
  filter(key == "AF") |> 
  group_by(key) %>%
  summarise(avg_effectSize = mean(effectSize, na.rm = TRUE)) # 0.3851843	

# df_raw %>% filter(key == "AF") |> pull(effectSize) |> mean()

# get 95% CI
df_raw %>%
  filter(key == "AF") |> 
  group_by(key) %>%
  summarise(avg_effectSize = mean(effectSize, na.rm = TRUE),
            lower = quantile(effectSize, 0.025, na.rm = TRUE),
            upper = quantile(effectSize, 0.975, na.rm = TRUE)) # 0.3851843, -0.6427598,	2.142308	for AF




# plot distribution for all keys
plt_4distributions <- df_raw |> 
  ggplot(aes(x = effectSize)) + 
  geom_density(aes(fill = key), alpha = 0.6, linewidth = 0.8) +  # Map fill to key
  facet_wrap(~key) + 
  geom_vline(xintercept = lower_quantile, color = "grey50", linetype = "dashed", linewidth = 0.8) +
  geom_vline(xintercept = upper_quantile, color = "grey50", linetype = "dashed", linewidth = 0.8) +
  scale_fill_scico_d(palette = "managua") +  # Use a Scico palette
  labs(
    x = "Effect Size", 
    y = "Density",
    title = "Distribution of effect sizes by study"
  ) +
  theme_classic() + 
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(), 
    strip.text = element_text(size = 12)
  )



# stack them
plt_density_plot_combined <- df_raw |> 
  ggplot(aes(x = effectSize, fill = key)) +  # Add fill for different keys
  geom_density(alpha = 0.5, linewidth = 0.8) +  # Adjust alpha and linewidth
  scale_fill_scico_d(palette = "managua") +  # Use a Scico palette
  geom_vline(xintercept = lower_quantile, color = "grey50", linetype = "dashed", linewidth = 0.8) +
  geom_vline(xintercept = upper_quantile, color = "grey50", linetype = "dashed", linewidth = 0.8) +
  labs(
    x = "Effect Size", 
    y = "Density",
    title = "Distribution of effect sizes by study",
    caption = "Dashed lines represent 95% confidence interval"
  ) +
  theme_classic() +  # Cleaner theme
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    panel.grid.minor = element_blank(),  # Remove minor grid lines
    legend.position = "bottom",  # Place legend at the bottom
    legend.title = element_blank()  # Remove legend title
  )




plt_4distributions
plt_density_plot_combined
```

# 4 histograms - no density plot

EDIT 18.11.2024: Own y-exis for each plot

EDIT 8.11.2024: instead of density plot, we need histograms
"Instead of the density plot we rather need the histograms of the 4 managements – exactly like the density plot but just the histograms.
-The axis texts and labels are too small when I tried to put it in the paper. We can make them bigger for readability."


```{r}
## 4 hist with free y
plt_4histograms <- df_raw |> 
  ggplot(aes(x = effectSize)) + 
  geom_histogram(aes(fill = key), alpha = 0.6, color = "black", bins = 40) +  # optimal number of bins?
  facet_wrap(~key, scales = "free_y") + 
  geom_vline(xintercept = lower_quantile, color = "grey50", linetype = "dashed", linewidth = 0.8) +
  geom_vline(xintercept = upper_quantile, color = "grey50", linetype = "dashed", linewidth = 0.8) +
  scale_fill_scico_d(palette = "managua") +  # Use a Scico palette
  labs(
    x = "Effect Size", 
    y = "Density",
    title = "Distribution of effect sizes by study"
  ) +
  theme_classic() + 
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 16),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(), 
    strip.text = element_text(size = 16)
  )
plt_4histograms

# NOTES: remove the "key" from the legend

```



# Sensitivity analysis

This takes 2min to run

```{r}

# Function for this
f_calculate_avg_effectSize_bykey <- function(key_to_be_included, obs_to_be_excluded){
  
  avg_effectSize <- 
    df_raw |> 
    filter(key == key_to_be_included) |> 
    mutate(ID = row_number()) |> # add ID
    filter(ID != obs_to_be_excluded) %>% 
    pull(effectSize) |>
    mean(na.rm=T) 
  
  # Create a data frame to store the result
  result <- tibble(
    key = key_to_be_included,
    excluded_obs = obs_to_be_excluded,
    avg_effectSize = avg_effectSize
  )
  
  return(result)
}

# try
f_calculate_avg_effectSize_bykey(key_to_be_included = "AF",  obs_to_be_excluded = 1) # 0.3842059	
f_calculate_avg_effectSize_bykey(key_to_be_included = "AF",  obs_to_be_excluded = 10) # 0.3855088	


# Apply this function to all keys (could be done with a function and lapply too)
avg_mean_in_AF <- map_df(
  unique(df_raw |> filter(key == "AF") |>  mutate(ID = row_number()) |>  pull(ID)), 
  ~f_calculate_avg_effectSize_bykey(key_to_be_included = "AF", obs_to_be_excluded = .x)
)

avg_mean_in_CC <- map_df(
  unique(df_raw |> filter(key == "CC") |>  mutate(ID = row_number()) |>  pull(ID)), 
  ~f_calculate_avg_effectSize_bykey(key_to_be_included = "CC", obs_to_be_excluded = .x)
)

avg_mean_in_NT <- map_df(
  unique(df_raw |> filter(key == "NT") |>  mutate(ID = row_number()) |>  pull(ID)), 
  ~f_calculate_avg_effectSize_bykey(key_to_be_included = "NT", obs_to_be_excluded = .x)
)

avg_mean_in_OF <- map_df(
  unique(df_raw |> filter(key == "OF") |>  mutate(ID = row_number()) |>  pull(ID)), 
  ~f_calculate_avg_effectSize_bykey(key_to_be_included = "OF", obs_to_be_excluded = .x)
)

# Combine
avg_mean_all <- bind_rows(avg_mean_in_AF, avg_mean_in_CC, avg_mean_in_NT, avg_mean_in_OF)


# quantiles for each key
df_quantiles <- avg_mean_all |> 
  group_by(key) |> 
  summarise(
    lower = quantile(avg_effectSize, 0.025),
    upper = quantile(avg_effectSize, 0.975)
  )




# ------------------------------------------------------------------------- Plot
plt_4_jacknifes <- ggplot(avg_mean_all, aes(x = excluded_obs, y = avg_effectSize)) +
  geom_point(aes(color = key), size = 1, alpha = 0.6) +
  geom_hline(data = df_quantiles, aes(yintercept = lower), linetype = "dashed", color = "black") +
  geom_hline(data = df_quantiles, aes(yintercept = upper), linetype = "dashed", color = "black") +
  facet_wrap(~key, scales = "free") +
  scale_color_scico_d(palette = "managua") +  
  labs(
    x = "Excluded observation ID",
    y = "Average effect size",
    title = "Sensitivity analysis by excluding one observation at a time"
  ) +
  theme_classic() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 16),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = 16),
    legend.position = "none"  
  )






# All together, no keys separated
# ---------------------------------------------------------------------------
# ---------------------------------------------------------------------------
f_avg_effectSize <- function(obs_to_be_excluded) {
  
  avg_effectSize <- 
    df_raw |> 
    mutate(ID = row_number()) |>
    filter(ID != obs_to_be_excluded) %>% 
    summarise(avg_effectSize = mean(effectSize, na.rm = TRUE)) %>% 
    pull(avg_effectSize)
  
  # Create a data frame to store the result
  result <- tibble(
    excluded_obs = obs_to_be_excluded,
    avg_effectSize = avg_effectSize
  )
  
  return(result)
}

f_avg_effectSize(1) # 0.05262971	

# Apply the function to all observations
avg_mean_in_all <- map_df(
  unique(df_raw |> mutate(ID = row_number()) |> pull(ID)), 
  ~f_avg_effectSize(.x)
)



# Plot
plt_1_jacknife <- 
  ggplot(avg_mean_in_all, aes(x = excluded_obs, y = avg_effectSize)) +
  geom_point(size = 1, alpha = 0.6, color = "#2C3E50") +  # Yhtenäinen väri kaikille pisteille
  geom_hline(yintercept = lower_quantile, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = upper_quantile, linetype = "dashed", color = "grey50") +
  labs(
    x = "Excluded Observation ID",
    y = "Average effect Size",
    title = "Excluding One observation at a time - average effect size"
  ) +
  theme_classic() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 16),
    panel.grid.minor = element_blank()
  )


plt_4_jacknifes
plt_1_jacknife
```

# Save

```{r}

# plt_4distributions
plt_4histograms
plt_density_plot_combined
plt_4_jacknifes


# Save as pdf files
ggsave(here("figures", "plt_density_plot_combined.pdf"), plt_density_plot_combined)
ggsave(here("figures", "plt_4histograms.pdf"), plt_4histograms)
ggsave(here("figures", "plt_4_jacknifes.pdf"), plt_4_jacknifes)




```




# Sensitivivty analysis - Jacknife (crops separated)

How much variable effectSize changes if we exclude one study at a time?
For example, when we exclude key == OF, how much is the average effectSize related to different crop species?

```{r}
# Example without function
# df_raw %>%
#   filter(key != "OF") %>%
#   group_by(Crop_Group) %>%
#   summarise(avg_effectSize = mean(effectSize, na.rm = TRUE))

# Function for this (drop when Crop_Group is NA)
f_calculate_avg_effectSize <- function(key_to_be_excluded){
  
  avg_effectSize <- 
    df_raw %>% 
    filter(key != key_to_be_excluded) %>% 
    group_by(Crop_Group) %>% 
    summarise(avg_effectSize = mean(effectSize, na.rm = TRUE)) %>% 
    drop_na() %>% 
    mutate(excluded_key = key_to_be_excluded) # Lisätään sarake, joka ilmoittaa poistetun keyn
  
  return(avg_effectSize)
}

# Example
f_calculate_avg_effectSize("OF")

# Run function for all keys
all_keys <- unique(df_raw$key)

# Sovelletaan funktiota kaikkiin key-arvoihin ja tallennetaan tulokset
results <- map_df(all_keys, f_calculate_avg_effectSize)
results

```



# Plot the results (initial, crop species separated)

This was later modified

```{r}
# Wide 
results_wide_arranged <- results %>%
  pivot_wider(names_from = Crop_Group, values_from = avg_effectSize) %>%
  arrange(excluded_key)

results_wide_arranged

# Long
results_long <- results_wide_arranged %>%
  pivot_longer(cols = -excluded_key, names_to = "Crop_Group", values_to = "avg_effectSize")



# Draw a combined plot 
# similarly as done below wiht plt_combined_arranged -- use facte_grid etc
library(scico)
plt_jacknife <-
  ggplot(results_long, 
         aes(x = factor(excluded_key, levels = unique(excluded_key)), 
             y = avg_effectSize, color = Crop_Group, group = Crop_Group)) +
  geom_line(linewidth = 1.2) +
  facet_grid(Crop_Group ~ ., 
             scales = "free_y")+#,
             #labeller = as_labeller(c("AF" = "AF", "CC" = "CC", "NT" = "NT"))) + # not necessary?
  theme_gray() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6),
        axis.text.y = element_text(size = 16),
        strip.text = element_text(size = 16),
        legend.position = "none") +
  scale_color_scico_d(palette = "roma") +
  labs(title = "Sensitivity analysis - Jacknife",
       x = "Excluded study",
       y = "Average effect size") 
 

plt_jacknife 
```












################## OLD CODE FOR REFERENCE - Daniel's study -- feel free to test
################## OLD CODE FOR REFERENCE - Daniel's study -- feel free to test

Code written by Johannes Piipponen

# Get and explore data

```{r}
# Load necessary libraries
library(tidyverse)
library(fixest)

# Read the data
df_raw <- read_csv("data_johannes.csv")

# Add a new column ln_gni as the natural logarithm of gni
df_raw <- df_raw %>% 
  mutate(ln_gni = log(gni))

# Explore data: number of countries, crops and years
length(unique(df_raw$country)) # 81 countries
unique(df_raw$year) %>% sort() # years 2000-2020



# Group data by country and count the number of cases
df_raw %>% 
  group_by(country) %>% 
  summarise(Number_of_cases = n()) %>% 
  arrange(Number_of_cases)

# Create a list of countries with more than 80 observations
#my_cntrylist_over80 <- 
my_countries_obs <-
  df_raw %>% 
  group_by(country) %>% 
  summarise(Number_of_cases = n()) %>% 
 # filter(Number_of_cases > 80) %>% 
  arrange(Number_of_cases)# %>%   pull(country)


```






# Sensitivity analysis

The following code performs a regression analysis using all available data and then explores the sensitivity of the model by excluding individual countries from the analysis.

When all countries are included in the model, it produces specific values for the intercept and beta1 coefficients. However, if we exclude a particular country (e.g., Finland), these coefficient values may change. The objective is to determine the magnitude of these changes.

The ultimate goal is to create a graph that visualizes how the regression results vary when we exclude one country at a time.

Explanation for f_exclude_countries function:
The f_exclude_countries function accepts the country to be excluded as an argument (country_tobe_excluded). It runs the feols model with the specified formula and filters the dataset so that the specified country is not included. After fitting the model, the function uses broom::tidy() to extract the regression results and selects the relevant columns (term, estimate, and p-value). Lastly, the function adds a new column, excluded_country, to indicate the country that was excluded in that particular model run.


```{r}
# Create a function to run feols excluding one country at a time
f_exclude_countries <- function(country_tobe_excluded) {
  feols(loss_percentage1 ~ ln_gni + agriShr + agriEmp + msch + 
           electricityRural + phoneSub + polStab + export + regionalName + as.factor(year),
         cluster = ~ regionalName, 
         data = df_raw %>% filter(country != country_tobe_excluded)) %>%
    broom::tidy() %>%
    dplyr::select(term, estimate, p.value) %>%
    mutate(excluded_country = country_tobe_excluded)
}

# Example function calls
f_exclude_countries("Angola")
f_exclude_countries("Burundi")

# Use the function for all countries
my_countries <- unique(df_raw$country)

# Apply the function to all countries and store the results
results <- map_df(my_countries, f_exclude_countries)

# Widen the results and rename p-value columns ---> not sure if this is needed
results_wider <- results %>%
  pivot_wider(names_from = term, values_from = c(estimate, p.value)) %>%
  rename(
    pval_intercept = `p.value_(Intercept)`,
    pval_ln_gni = `p.value_ln_gni`,
    pval_agriShr = `p.value_agriShr`,
    pval_agriEmp = `p.value_agriEmp`,
    pval_msch = `p.value_msch`,
    pval_electricityRural = `p.value_electricityRural`,
    pval_phoneSub = `p.value_phoneSub`,
    pval_polStab = `p.value_polStab`,
    pval_export = `p.value_export`
  )

results_wider
```


# Save the original regression

```{r}
library(flextable)
library(modelsummary)

model <-   feols(loss_percentage1 ~ ln_gni + agriShr + agriEmp + msch + 
                   electricityRural + phoneSub + polStab + export + regionalName + as.factor(year),
                 cluster = ~ regionalName, 
                 data = df_raw)



# Create the table for the feols model
model_table <- modelsummary(model, stars = T, output = "flextable")
model_table
# uncomment this
#save_as_docx(model_table, path = ("selected_results.docx"))
```


# Plot the results

# Sensitivity like originally but when number of cases is used to define X axis

```{r}
results_arranged <- map_df(unique(my_countries_obs$country), f_exclude_countries)

results_wider_arranged <- results_arranged %>%
  pivot_wider(names_from = term, values_from = c(estimate, p.value)) %>%
  rename(
    pval_intercept = `p.value_(Intercept)`,
    pval_ln_gni = `p.value_ln_gni`,
    pval_agriShr = `p.value_agriShr`,
    pval_agriEmp = `p.value_agriEmp`,
    pval_msch = `p.value_msch`,
    pval_electricityRural = `p.value_electricityRural`,
    pval_phoneSub = `p.value_phoneSub`,
    pval_polStab = `p.value_polStab`,
    pval_export = `p.value_export`
  )


results_long_arranged <- results_wider_arranged %>%
  pivot_longer(
    cols = starts_with("estimate_"),
    names_to = "variable",
    values_to = "value"
  ) %>% 
  dplyr::select(excluded_country, variable, value) %>%
  drop_na()



# Select only some variables
results_long_arranged <- results_long_arranged %>% 
  filter(variable %in% c("estimate_(Intercept)", "estimate_ln_gni",                                 
 "estimate_agriShr",                                  
  "estimate_agriEmp",                                  
  "estimate_msch",                                     
  "estimate_electricityRural",                         
  "estimate_phoneSub",                                 
  "estimate_polStab",                                  
  "estimate_export"   ))


# Draw a combined plot
library(scico)

plt_combined_arranged <- ggplot(results_long_arranged, aes(x = factor(excluded_country, levels = unique(excluded_country)), y = value, color = variable, group = variable)) +
  geom_line(size = 1.2) + 
  facet_grid(variable ~ .,
             scales = "free_y",
             labeller = as_labeller(c("estimate_(Intercept)" = "Intercept",
                                      estimate_ln_gni = "Log GNI",
                                      estimate_agriShr = "Agri Share",
                                      estimate_agriEmp = "Agri Emp",
                                      estimate_msch = "Mean Sch",
                                      estimate_electricityRural = "Elec Acc",
                                      estimate_phoneSub = "Phone Sub",
                                      estimate_polStab = "Pol Stab",
                                      estimate_export = "Exports"))) +
  theme_gray() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6),
        axis.text.y = element_text(size = 8),
        strip.text = element_text(size = 8),
        legend.position = "none") +
  scale_color_scico_d(palette = "roma") +
  labs(title = "Impact of excluding a country on regression output",
       x = "Excluded country countries with fewer observations         countries with more observations",
       y = "Estimate")


plt_combined_arranged

# uncomment this

# ggsave("sensitivity_plot_x_axis_based_on_N_obs.png", plot = plt_combined_arranged, dpi = 300, width = 300, height = 225, units = "mm")
# 
# ggsave("sensitivity_plot_x_axis_based_on_N_obs.pdf", plot = plt_combined_arranged)


```